Maintenance planning on French military aircraft operations

### 2018-02-19 R/roadef_presentation/roadef_presentation.Rmd
Added changes in ROADEF presentation after last test.

---
title: "Maintenance planning on French military aircraft operations"
author: "Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin"
date: "February 14, 2018"
output: 
  beamer_presentation:
    theme: "Madrid"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 2
    toc: true
nocite: '@*'
bibliography: MFMP.bib
---

@@

# Introduction

## Introduction

Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.

They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.

@@

Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have particular needs in terms of aircraft and time.

@@

```{r, echo=FALSE, out.width='100%'}
if (show_gantt){
    htmltools::HTML('<div style="font-size: 20px;width:100%">')
    gantt    
    htmltools::HTML("</div>")
} else {
    knitr::include_graphics('./../../img/calendar.png')
}

@@

## Problem (formally)

@@

## Relevance and origin

This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity.

In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan.

@@

* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and sustainability.
* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.

@@

## Model: sets

|Set|Description|
|-----|--------------------------------------------------------------------|
|    $\mathcal{I}$       |  resources. |
|    $\mathcal{T}$       |  time periods inside planning horizon. |
|    $\mathcal{J}$       |  tasks. |
|    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
|    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. |
|    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. |
|    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. |
|    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. |

## Model: main parameters

|Parameters|Description|
|--------------|-------------------------------------------------|
| $H_j$ |  amount of **resource time required** by task $j$. |
| $R_j$ |  number of **resources required** by task $j$. |
| $M$   |  **maintenance duration** in number of periods. |
| $E$   |  remaining **elapsed time** after a maintenance. |
| $H$   |  remaining **usage time** after a maintenance. |

## Model: more parameters

|Parameters|Description|
|--------------|----------------------------------------------------------------------|
| $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.|
| $D_t$              | number of resources to be assigned in total in period $t$. |
| $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. |
| $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. |
| $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. |
| $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. |

## Model: variables

| Variable   |  Description |
|-----------|------------------------------------------------------- |
| $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. |
| $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. |
| $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
| $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
| $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. |
| $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. |

@@

Requirements per task:
```{r out.width="250px", echo=FALSE}
knitr::include_graphics("./../../img/tasks_gantt.png")
```
\pause

Max maintenances:
```{r out.width="150px", echo=FALSE}
knitr::include_graphics("./../../img/num-maintenances.png")
```

@@

Visualization of accumulated time constraints for one resource.

```{r out.width="250px", echo=FALSE}
# img <- readPNG("./../../img/balance2.png")
#  grid.raster(img)
knitr::include_graphics("./../../img/balance2.png")
```

@@

Elapsed times are treated in the same way but consumption is always 1.
\begin{align}
& ret_{it} \leq ret_{it-1} + E m_{it} - 1 & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& ret_{i0} = Ret^{Init}_i
        & i \in \mathcal{I}\\
& ret_{it} \geq E m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& ret_{it} \in [0,E]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

@@

@@

Comparison of two objectives: max availability or maintenance capacity.
Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1
```{r out.width="200pt", echo=FALSE}
# img <- readPNG()
#  grid.raster(img)
knitr::include_graphics("./../../img/multiobjective.png")

Pareto diagram for instance id=I_5.

@@

* Analysis for all instances: not many pareto optimal points available.
* Between 2 and 4 for all instances
* The objectives are quite aligned one with the other.

@@

* **Try different objectives**: availability and maintenances.
* **Try more linent objectives**: relax the `max` in the objective.
* **Break symmetries**: clustering candidates, fixing end availabilities.
* **Add more constraints**: minimal duration of assignments, storage of aircraft.

# References
## References

>>>>

---
title: "Maintenance planning on French military aircraft operations"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
institute: "ISAE SUPAERO"
date: "February 22, 2018"
output: 
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 2
    toc: true
nocite: '@*'
bibliography: MFMP.bib
---

@@

<!-- # Introduction -->

<!-- ## Introduction -->

<!-- Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.  -->

<!-- They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours. -->

@@

Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have fixed start and end times and have particular needs in terms of aircraft and time.

@@

```{r, echo=FALSE, out.width='70%'}

if (show_gantt){
    htmltools::HTML('<div style="font-size: 20px;width:100%">')
    gantt
    htmltools::HTML("</div>")
} else {
    knitr::include_graphics('./../../img/calendar.png')
}

@@

## Problem

@@

<!-- ## Relevance and origin -->

<!-- This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity. -->

<!-- In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan. -->

@@

* FMP: Flight and Maintenance Planning problem.
* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.

@@

<!-- ## Model: sets -->

<!-- |Set|Description| -->
<!-- |-----|--------------------------------------------------------------------| -->
<!-- |    $\mathcal{I}$       |  resources. | -->
<!-- |    $\mathcal{T}$       |  time periods inside planning horizon. | -->
<!-- |    $\mathcal{J}$       |  tasks. | -->
<!-- |    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. | -->
<!-- |    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. | -->
<!-- |    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. | -->
<!-- |    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. | -->
<!-- |    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. | -->

<!-- ## Model: main parameters -->

<!-- |Parameters|Description| -->
<!-- |--------------|-------------------------------------------------| -->
<!-- | $H_j$ |  amount of **resource time required** by task $j$. | -->
<!-- | $R_j$ |  number of **resources required** by task $j$. | -->
<!-- | $M$   |  **maintenance duration** in number of periods. | -->
<!-- | $E$   |  remaining **elapsed time** after a maintenance. | -->
<!-- | $H$   |  remaining **usage time** after a maintenance. | -->

<!-- ## Model: more parameters -->

<!-- |Parameters|Description| -->
<!-- |--------------|----------------------------------------------------------------------| -->
<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->

<!-- ## Model: variables -->

<!-- | Variable   |  Description | -->
<!-- |-----------|------------------------------------------------------- | -->
<!-- | $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. | -->
<!-- | $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. | -->
<!-- | $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
<!-- | $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
<!-- | $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. | -->
<!-- | $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. | -->

<!-- ## Model: main constraints -->

<!-- Requirements per task: -->
<!-- ```{r out.width="250px", echo=FALSE} -->
<!-- knitr::include_graphics("./../../img/tasks_gantt.png") -->
<!-- ``` -->
<!-- \pause -->

<!-- Max maintenances: -->
<!-- ```{r out.width="150px", echo=FALSE} -->
<!-- knitr::include_graphics("./../../img/num-maintenances.png") -->
<!-- ``` -->

<!-- ## Model: main constraints -->

## Model: objective function

\begin{align}
    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
\end{align}

@@

@@

<!-- ## Model: flow constraints -->

<!-- Visualization of accumulated time constraints for one resource. -->

<!-- ```{r out.width="250px", echo=FALSE} -->
<!-- # img <- readPNG("./../../img/balance2.png") -->
<!-- #  grid.raster(img) -->
<!-- knitr::include_graphics("./../../img/balance2.png") -->
<!-- ``` -->

@@

Elapsed times are treated in the same way but consumption is expressed in time periods.
\begin{align}
& ret_{it} \leq ret_{it-1} + E m_{it} - 1 & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& ret_{i0} = Ret^{Init}_i
        & i \in \mathcal{I}\\
& ret_{it} \geq E m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& ret_{it} \in [0,E]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

@@

## Sample solution progress

The analysis of the solution progress for all instances shows that:

* Initial solutions are quite far from the linear relaxation.
* Branching does not improve significantly the lower bound.

\pause
```{r out.width="150pt", echo=FALSE}
knitr::include_graphics("./../../img/progress.png.png")
```

*(Progress diagram for instance id=I_7).*

@@

Comparison of two objectives: max availability or maintenance capacity.
Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1.
\pause
```{r out.width="180pt", echo=FALSE}
knitr::include_graphics("./../../img/multiobjective.png")
```

*(Pareto diagram for instance id=I_5).*

@@

* Analysis for all instances: not many pareto optimal points available (between two and four).
* The objectives are quite aligned one with the other.

@@

* **Try different objective formulations**: availability - maintenability - future state.
* **Try more linent objectives**: relax the `max` in the objective.
* **Break symmetries**: clustering candidates, fixing end availabilities.
* **Add more constraints**: minimal duration of assignments, storage of aircraft.

# References
## References / Questions?
